Empirical Analysis of the Performance Bottleneck in Graph Neural Network Training

### 2020-08-29 CCPE/Mauscript/wileyNJD-AMA.tex
更新CCPE版本的前言，同时重命名了FGCS.

\author[1]{Zhaokang Wang*}
\author[2,3]{Yunpan Wang, Chunfeng Yuan}
\author[3]{Rong Gu, Yihua Huang}

\authormark{AUTHOR ONE \textsc{et al}}

\address[1]{\orgdiv{Pasa Lab}, \orgname{Nanjing University}, \orgaddress{\state{JiangSu}, \country{China}}}
\address[2]{\orgdiv{Pasa Lab}, \orgname{Nanjing University}, \orgaddress{\state{JiangSu}, \country{China}}}
\address[3]{\orgdiv{Pasa Lab}, \orgname{Nanjing University}, \orgaddress{\state{JiangSu}, \country{China}}}

\corres{*Rong Gu, *Yihua Huang. \email{\{gurong, yhuang\}@nju.edu.cn}}

\presentaddress{No.163, Xianlin Road, Qixia District, Nanjing City, Jiangsu Province, China}

>>>>

\author[1]{Zhaokang Wang, Yunpan Wang, Chunfeng Yuan, Rong Gu*, Yihua Huang*}

\authormark{Wang \textsc{et al}}

\address{\orgdiv{State Key Laboratory for Novel Software Technology}, \orgdiv{Department of Computer Science and Technology}, \orgname{Nanjing University}, \orgaddress{\state{Jiangsu}, \country{China}}}

\corres{Rong Gu* and Yihua Huang* with equal contribution. \\ \email{\{gurong, yhuang\}@nju.edu.cn}}

\presentaddress{No.163 Xianlin Avenue,  Nanjing 210023, Jiangsu, China}

### 2020-08-29 FGCS/sec1-introduction.tex
更新CCPE版本的前言，同时重命名了FGCS.

\section{Introduction}
In recent years, the graph neural network (GNN) becomes a hot research topic in the field of artificial intelligence.
Many GNNs~\cite{kipf2017_gcn, defferrad2016_chebnet, li2018_agcn,li2015_ggnn, hamilton2017_graphsage, huang2018_gat, zhang2018_gaan} are proposed.
They can learn the representation of vertices/edges in a graph from its topology and the original feature vectors in an \emph{end-to-end} manner.
The powerful expression ability of GNNs makes them achieve good accuracy in many graph analysis tasks \cite{zhou2018_gnn_review, zhang2018_gnn_survey, comprehensive-survey-wu-2020}, like node classification, link prediction, and graph classification.

>>>>

\section{Introduction}

In recent years, the graph neural network (GNN) becomes a hot research topic in the field of artificial intelligence.
Many GNNs \cite{kipf2017_gcn, defferrad2016_chebnet, li2018_agcn,li2015_ggnn, hamilton2017_graphsage, huang2018_gat, zhang2018_gaan} are proposed.
They can learn the representation of vertices/edges in a graph from its topology and the original feature vectors in an \emph{end-to-end} manner.
The powerful expression ability of GNNs makes them achieve good accuracy in many graph analysis tasks \cite{zhou2018_gnn_review, zhang2018_gnn_survey, comprehensive-survey-wu-2020}, like node classification, link prediction, and graph classification.

### 2020-08-29 FGCS/sec6-related-work.tex
更新CCPE版本的前言，同时重命名了FGCS.

\paragraph{Survey of GNNs}
\cite{zhou2018_gnn_review,zhang2018_gnn_survey,comprehensive-survey-wu-2020} survey the existing graph neural network models and classify them from an algorithmic view.
They summarize the parallels and differences between the architecture of different GNN models.
The typical applications of GNNs are also briefly introduced.
Those surveys focus on comparing the existing GNNs theoretically not empirically.

\paragraph{Evaluation of GNNs}
Shchur et al. \cite{shchur2018_pitfall_of_gnn} evaluate the accuracy of popular GNN models on the node classification task.
Dwived et al. \cite{dwivedi2020_benchmark_of_gnn} further compare the accuracy of popular GNNs fairly in a controlled environment.
Hu et al. \cite{hu2020_open_graph_benchmark} propose the open graph benchmark that provides a group of datasets and a standard evaluation workflow.
The benchmark makes comparisons between GNNs easily and fairly.
Those model evaluation efforts focus on evaluating the accuracy of different GNNs.
They provide insightful suggestions to improve accuracy.

From the efficiency aspect, Yan et al. \cite{yan2020_characterizing_gcn} compare the performance characteristics of graph convolutional networks, typical graph processing (like PageRank), and MLP-based neural network on GPU.
They provide optimization guidelines for the software and the hardware sides.
Zhang et al. \cite{zhang2020_analysis_neugraph} compare the architectural characteristics of the GNN inference on GPU under the unified SAGA-NN \cite{ma2019_neugraph} programming model.
They find that GNN inference has no fixed performance bottleneck and all components deserve to optimize.
These two efforts focus on the \emph{inference} phase of GNNs and they investigate the potential optimizations from an architectural view.
In this work, our target is to find out the performance bottleneck in the training phase from a system view.
We consider the performance bottleneck in both time and memory usage.
We also evaluate the effects of the sampling techniques.
Our work and \cite{yan2020_characterizing_gcn, zhang2020_analysis_neugraph} form a complementary study on the efficiency issue of GNNs.

>>>>

\paragraph{Survey of GNNs}
Zhou et al.\cite{zhou2018_gnn_review}, Zhang et al.\cite{zhang2018_gnn_survey} and Wu et al.\cite{comprehensive-survey-wu-2020} survey the existing graph neural networks and classify them from an algorithmic view.
They summarize the similarities and differences between the architectures of different GNNs.
The typical applications of GNNs are also briefly introduced.
Those surveys focus on comparing the existing GNNs theoretically, not empirically.

\paragraph{Evaluation of GNNs}
Shchur et al. \cite{shchur2018_pitfall_of_gnn} evaluate the accuracy of popular GNNs on the node classification task.
Dwived et al. \cite{dwivedi2020_benchmark_of_gnn} further compare the accuracy of popular GNNs fairly in a controlled environment.
Hu et al. \cite{hu2020_open_graph_benchmark} propose the open graph benchmark that provides a standard datasets and a standard evaluation workflow.
The benchmark makes comparisons between GNNs easily and fairly.
Those model evaluation efforts focus on evaluating the accuracy of different GNNs.
They provide insightful suggestions to improve accuracy.

From the efficiency aspect, Yan et al. \cite{yan2020_characterizing_gcn} compare the performance characteristics of graph convolutional networks, typical graph processing kernels(like PageRank), and the MLP-based neural networks on GPU.
They provide optimization guidelines for both the software and the hardware developers.
Zhang et al. \cite{zhang2020_analysis_neugraph} analyze the architectural characteristics of the GNN inference on GPU under SAGA-NN \cite{ma2019_neugraph} model.
They find that the GNN inference has no fixed performance bottleneck and all components deserve to optimize.
These two efforts focus on the \emph{inference} phase of GNNs and they investigate the potential optimizations mainly from an architectural view.
In this work, our target is to find out the performance bottleneck in the training phase from a system view.
We consider the performance bottleneck in both time and memory usage.
We also evaluate the effects of the sampling techniques.
Our work and \cite{yan2020_characterizing_gcn, zhang2020_analysis_neugraph} form a complementary study on the efficiency issue of GNNs.
